---
title: "NYC Air Pollution"
author: "Lizzy Gibson"
date: "April 22, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
library(forcats)
library(rvest)
library(httr)
library(janitor)
library(leaflet)
library(reshape2)
library(ggmap)
library(maps)
library(mapdata)
```

# Clean Data

Daily EPA Air Quality System Data for 2015 from: https://aqs.epa.gov/aqsweb/airdata/download_files.html.

## Download 

### Criteria Gases

```{r download_2015}
tmp <- tempfile()   
GET("https://aqs.epa.gov/aqsweb/airdata/daily_44201_2015.zip", write_disk(tmp))
ozone <- unzip(tmp) %>% read_csv() %>% clean_names() %>% 
  filter(state_name == "New York" & county_name %in% 
           c("Queens", "Kings", "New York", "Bronx", "Richmond") & poc == 1)

tmp2 <- tempfile()   
GET("https://aqs.epa.gov/aqsweb/airdata/daily_42401_2015.zip", write_disk(tmp2))
so2 <- unzip(tmp2) %>% read_csv() %>% clean_names() %>% 
  filter(state_name == "New York" & county_name %in% 
           c("Queens", "Kings", "New York", "Bronx", "Richmond") & poc == 1) %>% 
  mutate(parameter_name = fct_recode(parameter_name, 
                                     "SO2" = "Sulfur dioxide"))

tmp3 <- tempfile()   
GET("https://aqs.epa.gov/aqsweb/airdata/daily_42101_2015.zip", write_disk(tmp3))
co <- unzip(tmp3) %>% read_csv() %>% clean_names() %>% 
  filter(state_name == "New York" & county_name %in% 
           c("Queens", "Kings", "New York", "Bronx", "Richmond") & poc == 1) %>%
  mutate(parameter_name = fct_recode(parameter_name, 
                                     "CO" = "Carbon monoxide"))

tmp4 <- tempfile()   
GET("https://aqs.epa.gov/aqsweb/airdata/daily_42602_2015.zip", write_disk(tmp4))
no2 <- unzip(tmp4) %>% read_csv() %>% clean_names() %>% 
  filter(state_name == "New York" & county_name %in% 
           c("Queens", "Kings", "New York", "Bronx", "Richmond") & poc == 1) %>% 
  mutate(parameter_name = fct_recode(parameter_name, 
                                     "NO2" = "Nitrogen dioxide (NO2)"))
```

### Particulates

Do not include PM10 or non-FRM (non-federal reference method).

```{r download_particulates}
tmp11 <- tempfile()   
GET("https://aqs.epa.gov/aqsweb/airdata/daily_88101_2015.zip", write_disk(tmp11))
pm_25_FRM_FEM_mass <- unzip(tmp11) %>% read_csv() %>% clean_names() %>% 
  filter(state_name == "New York" & county_name %in% 
           c("Queens", "Kings", "New York", "Bronx", "Richmond") & poc == 1) %>% 
  mutate(parameter_name = fct_recode(parameter_name, 
                                     "PM2.5" = "PM2.5 - Local Conditions"))

tmp44 <- tempfile()   
GET("https://aqs.epa.gov/aqsweb/airdata/daily_SPEC_2015.zip", write_disk(tmp44))
pm_25_spec <- unzip(tmp44) %>% read_csv() %>% clean_names() %>% 
  filter(state_name == "New York" & county_name %in% 
           c("Queens", "Kings", "New York", "Bronx", "Richmond") & 
           parameter_name %in% c("Sodium PM2.5 LC",
                                             "Aluminum PM2.5 LC",
                                             "Barium PM2.5 LC",
                                             "Bromine PM2.5 LC",
                                             "Calcium PM2.5 LC",
                                             "Chlorine PM2.5 LC",
                                             "Chromium PM2.5 LC",
                                             "Copper PM2.5 LC",
                                             "Iron PM2.5 LC",
                                             "Lead PM2.5 LC",
                                             "Manganese PM2.5 LC",
                                             "Nickel PM2.5 LC",
                                             "Selenium PM2.5 LC",
                                             "Silicon PM2.5 LC",
                                             "Sulfur PM2.5 LC",
                                             "Titanium PM2.5 LC",
                                             "Vanadium PM2.5 LC",
                                             "Zinc PM2.5 LC",
                                             "Black Carbon PM2.5 at 880 nm"))
table(pm_25_spec$poc)
```

### Meteorological

Do not include wind or barometric pressure.

```{r download_meteor}
tmp222 <- tempfile()   
GET("https://aqs.epa.gov/aqsweb/airdata/daily_TEMP_2015.zip", write_disk(tmp222))
temp <- unzip(tmp222) %>% read_csv() %>% clean_names() %>% 
  filter(state_name == "New York" & county_name %in% 
           c("Queens", "Kings", "New York", "Bronx", "Richmond") & poc == 1) %>% 
  mutate(parameter_name = fct_recode(parameter_name, 
                                     "Temp" = "Outdoor Temperature"))

tmp444 <- tempfile()   
GET("https://aqs.epa.gov/aqsweb/airdata/daily_RH_DP_2015.zip", write_disk(tmp444))
rh_dewpoint <- unzip(tmp444) %>% read_csv() %>% clean_names() %>% 
  filter(state_name == "New York" & county_name %in% 
           c("Queens", "Kings", "New York", "Bronx", "Richmond") & poc == 1) %>% 
  mutate(parameter_name = fct_recode(parameter_name, 
                                     "Humidity" = "Relative Humidity"))
```

### Toxics, Precursors, and Lead

Do not include toxics, precursors, or lead.

## Merge

Include criteria gases, PM2.5, PM2.5 species, temp, and relative humidity.

Restrict to NYC counties, POC = 1, and PM2.5 speciation from Kioumourtzoglou 2014 (https://www.nature.com/articles/jes20147#supplementary-information).

BC  
Na  
Al  
Si  
S  
Cl  
Ca  
Ti  
V  
Cr  
Mn  
Fe  
Ni  
Cu  
Zn  
Se  
Br  
Ba  
Pb  

```{r merge}
meteor <- rbind(rh_dewpoint, temp)
partic <- rbind(pm_25_FRM_FEM_mass, pm_25_spec)
criteria <- rbind(rbind(rbind(ozone, so2), co), no2)

all <- rbind(rbind(criteria, meteor), partic) %>% 
  select(date_local, latitude, longitude, parameter_name, arithmetic_mean) %>% 
  mutate(parameter_name = fct_recode(parameter_name, 
                                     "Sodium" = "Sodium PM2.5 LC",
                                             "Aluminum" = "Aluminum PM2.5 LC",
                                             "Barium" = "Barium PM2.5 LC",
                                             "Bromine" = "Bromine PM2.5 LC",
                                             "Calcium" = "Calcium PM2.5 LC",
                                             "Chlorine" = "Chlorine PM2.5 LC",
                                             "Chromium" = "Chromium PM2.5 LC",
                                             "Copper" = "Copper PM2.5 LC",
                                             "Iron" = "Iron PM2.5 LC",
                                             "Lead" = "Lead PM2.5 LC",
                                             "Manganese" = "Manganese PM2.5 LC",
                                             "Nickel" = "Nickel PM2.5 LC",
                                             "Selenium" = "Selenium PM2.5 LC",
                                             "Silicon" = "Silicon PM2.5 LC",
                                             "Sulfur" = "Sulfur PM2.5 LC",
                                             "Titanium" = "Titanium PM2.5 LC",
                                             "Vanadium" = "Vanadium PM2.5 LC",
                                             "Zinc" = "Zinc PM2.5 LC",
                                             "Black_Carbon" = "Black Carbon PM2.5 at 880 nm"))

aqs <- all %>%
  rename(date = date_local) %>% 
  group_by(date, parameter_name) %>% 
  summarize(mean = mean(arithmetic_mean, na.rm = TRUE)) %>% 
  spread(parameter_name, mean) %>% 
  filter(!is.na(Aluminum))

write_csv(all, "all_long.csv")
write_csv(aqs, "aqs_2015.csv")
```

Restrict to every 3 days for PM2.5 species.

## Monitor Stations

```{r map}
all %>% select(latitude, longitude) %>% unique() %>% 
  leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addCircles(~longitude, ~latitude, color = "red")

nyc_map <- get_map(location = c(lon = -74.00, lat = 40.71), maptype = "terrain", zoom = 11)

ggmap(nyc_map) + geom_point(data = all, aes(x = longitude, y = latitude), color = "red") + 
  labs(title = "Map of Monitor Stations across NYC", x = "Longitude", y = "Latitude")
```

