---
title: "Unsupervised Analysis"
author: "Lizzy Gibson"
date: "April 29, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
library(forcats)
library(rvest)
library(httr)
library(janitor)
library(leaflet)
library(reshape2)
library(ISLR)
library(ggfortify)
```

## PCA

A low-dimensional representation of the data that captures as much of the information as possible. The idea is that each of the n observations lives in p-dimensional space, but not all of these dimensions are equally interesting. PCA seeks a small number of dimensions that are as interesting as possible, where the concept of interesting is measured by the amount that the observations vary along each dimension. Each of the dimensions found by PCA is a linear combination of the p features. The first two principal components of a data set span the plane that is closest to the n observations, in terms of average squared Euclidean distance.

```{r data}
#remove outcome and confounders from PCA matrix
unsup <- data %>% select(-resp, -day_of_week, -outdoor_temperature, -relative_humidity, -date)
```

```{r}
pr.out <- prcomp(~ . , unsup, scale = TRUE, na.action = na.omit)
names(pr.out)

pc_out_plot <- as.tibble(pr.out$x) %>% 
  mutate(date = data[1:108,]$date) %>% 
  mutate(day_of_week = data[1:108,]$day_of_week) %>% 
  mutate(month = months(date)) 

pc_out_load <- as.tibble(pr.out$rotation)

pr.out$rotation
pr.var = pr.out$sdev^2
pr.var
# variance explained by each PC

pve = pr.var/sum(pr.var)
pve
# Percent Variance Explained

percentage <- round(pve * 100, 2)
percentage <- str_c(colnames(pr.out$x), " (", as.character(percentage), "%", ")")

#Scree plot
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", ylim = c(0,1), type = 'b')
#Elbow is at 3rd PC, meaning that the first 3 components explain most of the data

plot(cumsum(pve), xlab = "Principal Component", 
     ylab = "Cumulative Proportion of Variance Explained", ylim = c(0,1), type = 'b')

#heatmap of pollutant loadings on PC
autoplot(scale(pr.out$rotation))

pr.out$center
pr.out$scale
pr.out$rotation
# The kth column is the kth loading vector (phi)

dim(pr.out$x)
head(pr.out$x)
# The kth column is the kth PC score vector (z)

biplot(pr.out, scale = 0)
# scale=0: the arrows are scaled to represent the loadings
autoplot(pr.out, loadings = TRUE) + theme_bw()

ggplot(pc_out_plot) + geom_point(aes(x = PC1, y = PC2, color = month)) + 
  xlab(percentage[1]) + 
  ylab(percentage[2]) +
  theme_bw()

ggplot(pc_out_plot, aes(x = PC1, y = PC2, color = day_of_week)) + xlab(percentage[1]) + 
  ylab(percentage[2]) +
  geom_point() + theme_bw()

```

The principal component score vectors have length n = 122 for 122 days, and the principal component loading vectors have length p = 23 for 23 criteria gases, particulates, and PM species measured. PCA was performed after standardizing each variable to have mean zero and standard deviation one.

months -- seasons
days of week -- week/end
